---
output: 
  bookdown::pdf_document2:
    toc: false
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
title: "STP598-Assignment 1"
author: 
- Hao Wang 
- hwang306
date: '`r format(Sys.Date(), "%B %d, %Y")`'
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
spacing: single
papersize: letter
bibliography: D:/Dropbox/bib/Haowang.bib
biblio-style: apsr
citecolor: blue
header-includes: \usepackage{graphicx, longtable, float, subfigure}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Question 1

Load the data first, with a brief inspection.

```{r, echo=TRUE, include=TRUE}
mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Computational-Stats/master/Assignment1/HtVol.csv", header = TRUE)
summary(mydata)
```

## 1.a Predictive model

I build a simple linear model first

$$\hat{HtVol} =  \beta_0 + \beta_{1}Male + \beta_{2}Age + \beta_{3}Ht + \beta_{4}Wt$$

Use `lm` function to obtain the LS coefficients

```{r lm}
lm <- lm(HtVol ~ Male + Age + Ht + Wt, data = mydata)
summary(lm)
```

With the `lm` function, the fitting line is

$$\hat{HtVol} =  -263.67 + 41.34*Male -0.34*Age + 3.38*Ht + 0.68*Wt$$

I use package 'L1pack' to perform least absolute deviation regression

Unlike Least Squares, Least Absolute Deviation minimize

$$ S = \sum_{i =1}^{n}|y_i - f(x_i)|$$

```{r LAD, echo=TRUE, include=TRUE}
library("L1pack")
lad  <-  lad(HtVol ~ Male + Age + Ht + Wt, data = mydata)
summary(lad)
```

The fitting line is 

$$\hat{HtVol} =  -213.76 + 48.17*Male + 0.18*Age + 2.99*Ht + 4.41*Wt$$

## 1.b 
Steps are very similar, just change the explainatory variables

```{r}
lm <- lm(HtVol ~ Male + Age + BMI + BSA, data = mydata)
summary(lm)
```

The linear fitting is 

$$\hat{HtVol} = -121.95 + 37*Male -0.67* Age - 5.30 * BMI + 590 * BSA$$

For the LAD method:

```{r}
lad <- lad(HtVol ~ Male + Age + BMI + BSA, data = mydata)
summary(lad)
```

$$\hat{HtVol} = -40.99 + 37.87*Male -0.22* Age - 6.66 * BMI + 501.57 * BSA$$

