---
output: 
  bookdown::pdf_document2:
    toc: false
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
title: "STP598-Assignment 4"
author: 
- Hao Wang 
- hwang306
date: '`r format(Sys.Date(), "%B %d, %Y")`'
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
spacing: single
papersize: letter
citecolor: blue
header-includes: \usepackage{graphicx, longtable, float, subfigure}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

Create an R function named RLMstep that performs (hybrid) stepwise model selection based on the method illustrated in p. 11 of the article entitled 'Robust Stepwise Regression'. The function should be based on the existing rlm function; use the default settings. The function should take as input:
1. a column vector y (response)
2. a matrix X of predictors (predictors should be in columns)

Your function should report: a) the predictors that are included in the final
robust regression model, and b) coefficient estimates, and their BCa
confidence intervals


The hybrid method of stepwise regression considers both forward and backwards directions. Similar to forward stepwise regression, in a hybrid model, variables are added sequentially. However, after the new variable is added, this method also removes variables that are no longer statistically significant in the regression. The hybrid method mimics the power of best subset regression, but also has the computational advantages of forward and backwards stepwise regression.

I use the `step` function to perform a hybrid approach of regression. `step` function considers the AIC criteria for weighting the choices, at each step an add or drop will be performed to minimize the AIC score.

The robust regression function is taken from function `rlm`.

```{r}
library(MASS)
library(step)

mydata <- read.csv("https://raw.githubusercontent.com/haowang666/Computational-Stats/master/Assignment1/HtVol.csv", header = TRUE)
summary(mydata)

rlm <- rlm(HtVol ~ I(Male^2)+ I(CT^2) + I(Age^2) + I(Ht^2) +I(Wt^2) + I(BMI^2) + I(BSA^2) + .^2, psi = psi.huber, data = mydata, init = "ls")

rlm2 <- rlm(HtVol ~ .^2, data = mydata)

null <- rlm(HtVol ~ 1, data = mydata)
full <- rlm(HtVol ~ .^2, data = mydata)

rlm1 <- rlm(HtVol ~ Male + Age, data = mydata)
rlm2 <- rlm(HtVol ~ Male, data = mydata)

step = step(null , scope = list( lower = null , upper = full ),
             direction = c("forward"))

```




```{r}
library(wle) #This package reflects the paper Agostinelli 2002

data(selection)

result <- wle.stepwise(ydata~xdata, num.sol = 3,
min.weight=0.8, type="Stepwise", method="WLS")

summary(result)

library(MASS)
RLMstep <- function(y, x, boot){
  x <- cbind(1, x) #add intercept
  p <- ncol(x) # maximum number of predictors
  model.in <- NA # model building for stepwise regression. the initial status is null
  stepwise <- wle::wle.stepwise(y ~ x, num.sol = 3, min.weight = 0.8, type = "Stepwise", method = "WLS") #Fuction taken from the correspoding package
  col.in <- stepwise$wstep #for selecting columns
  for (i in 1:p) {
    index <- col.in[[i]]
    keep <- (index == 1)
    if (keep == TRUE) {
    model.in <- cbind(model.in, x[, i]) }
  }
  rlm <- rlm(y ~ model.in)
}
library(wle)
library(MASS)
data(selection)
x <- RLMstep(ydata, xdata, 3)   

data(selection)
result <- wle.stepwise(ydata~xdata, num.sol = 3,
min.weight=0.8, type="Stepwise", method="WLS")


x <- result$wstep

model.in <- NULL
xdata <- cbind(1, xdata)
keep <- numeric(4)

for (i in 1:4) {
    index <- x[[i]]
    keep[i] <- (index == 1)
    if (keep[i] == TRUE) {
    model.in <- cbind(model.in, xdata[, i]) }
  }

```





# Question 2

Compare your function versus the stepAIC function (direction= both") using the HtVol data from the first assignment.
The set of possible predictors includes Male; Age; Ht; Wt; BMI; BSA,
all quadratic terms and all bivariate interactions.

# Question 3

Discuss your findings; create a few figures that convey useful info wrt
your results.

